---
title: "Advanced Bioinformatics 2025 assessment"
author: "m2407447"
output: html_document
---
## Task 1

In this task, I will use the `sum()` function and the `:` operator in R to calculate the sum of all integers from 5 to 55.  
The `:` operator generates a sequence of numbers (in this case, 5 to 55), and `sum()` adds them together. In this case the sum is 1530

```{r}
sum(5:55)
```
## Task 2

In this task, I define a function called `sumfun()` that takes a single input parameter `n`, and calculates the sum of all integers from 5 to `n` using the `sum()` function and the `:` operator. I then use the function to compute the sum for `n = 10`, `n = 20`, and `n = 100`.

```{r}
# Define the function
sumfun <- function(n) {
  sum(5:n)
}
```

```{r}
# Use the function with different values of n
sumfun(10)
sumfun(20)
sumfun(100)

```
## Task 3

In this task, I generate the first 12 numbers in the Fibonacci sequence using a `for` loop in R. The Fibonacci sequence is defined such that each number is the sum of the two preceding ones, starting from 1 and 1.

```{r}
# Create a numeric vector to store Fibonacci numbers
fib <- numeric(12)

# Set the first two numbers
fib[1] <- 1
fib[2] <- 1

# Use a for loop to calculate the next numbers
for (i in 3:12) {
  fib[i] <- fib[i - 1] + fib[i - 2]
}

# Print the result
fib
```
## Task 4

In this task, I use the `ggplot2` package to create a boxplot of miles per gallon (`mpg`) as a function of the number of gears (`gear`) in the `mtcars` dataset. The `fill` aesthetic is used to colour the boxes based on the number of gears.

```{r, message=FALSE, warning=FALSE}
# Load ggplot2
library(ggplot2)

# Create the boxplot
ggplot(mtcars, aes(x = factor(gear), y = mpg, fill = factor(gear))) +
  geom_boxplot() +
  labs(
    title = "Miles per Gallon by Number of Gears",
    x = "Number of Gears",
    y = "Miles per Gallon (mpg)",
    fill = "Gears"
  ) +
  theme_minimal()
```
## Task 5

In this task, I fit a linear model to explore the relationship between a car's speed (`speed`) and its stopping distance (`dist`) using the `cars` dataset. The `lm()` function is used to compute the regression line.

### Step 1: Fit the linear model
```{r}
# Fit the linear model
model <- lm(dist ~ speed, data = cars)

# Show the model summary
summary(model)
```

### Step 2: Interpret the output

- **Fitted slope (speed coefficient)**: This tells us how much the stopping distance increases for every 1 mph increase in speed.
- **Fitted intercept**: This is the estimated stopping distance when speed = 0 mph.
- **Standard errors**: These represent the variability (uncertainty) in the slope and intercept estimates.

### Step 3: Units

- `speed` is measured in **miles per hour (mph)**  
- `dist` is measured in **feet**



